A SAT-Based Counterexample Guided Method for Unbounded Synthesis

### 2016-02-25 paper.tex
minor
Loop optimisation

An abstraction of the game restricts actions available to one of the players.
Specifically, we consider abstractions represented as trees of actions,
referred to as \emph{abstract game trees}.  Figure~\ref{fig:agt} shows an
example abstract game tree for the controller (abstract game trees for the
environment are similar).  In the abstract game, the controller is required to
pick actions from the tree, starting from the root node.  After reaching a
leaf, it continues playing unrestricted.  The tree in Figure~\ref{fig:agt}
restricts the initial controller action to the set $\{a, d\}$.  After choosing
action $a$, the controller is required to play action $b$ in the next round.
It then reaches a leaf of the tree and continues playing unrestricted.  The
root of the tree is annotated by the initial state $s$ of the abstract game and
the bound $k$ on the number of rounds.  We denote $\textsc{nodes}(T)$ the set
of all nodes of a tree $T$, $\textsc{leaves}(T)$ the subset of leaf nodes.  For
edge $e$, $\textsc{action}(e)$ is the action that labels the edge, and for node
$n$, $\textsc{height}(n)$ is the maximal distance from n to a leaf. For node
$n$ of the tree, $\textsc{succ}(n)$ is the set of pairs $\langle e, n' \rangle$
where $n'$ is a child node of $n$ and $e$ is the edge connecting $n$ and $n'$.

@@

\begin{algorithm}
    \begin{algorithmic}
        \Function{checkBound}{$k$}
            \State $r \gets $ \Call{solveAbstract}{$\texttt{env}, I, k, \emptyset$}
            \IIf{$r \neq \emptyset$} \Return $r$ \EndIIf
            \State $s' \gets I$
            \For{$x \in X$}
            \State $r \gets$ \Call{solveAbstract}{$\texttt{env}, s' \setminus \{x\}, k, \emptyset$} 
                \IIf{$r = \emptyset$} $s' \gets s' \setminus \{x\}$ \EndIIf \Comment{Removes the assignment to $x$ from $s'$}
            \EndFor
            \State \Return $\emptyset$
        \EndFunction
    \end{algorithmic}
    \caption{Generalise $I$ optimisation}
    \label{alg:opt1}
\end{algorithm}

>>>>

An abstraction of the game restricts actions available to one of the players.
Specifically, we consider abstractions represented as trees of actions,
referred to as \emph{abstract game trees}.  Figure~\ref{fig:agt} shows an
example abstract game tree for the controller (abstract game trees for the
environment are similar).  In the abstract game, the controller is required to
pick actions from the tree, starting from the root node.  After reaching a
leaf, it continues playing unrestricted.  The tree in Figure~\ref{fig:agt}
restricts the initial controller action to the set $\{a, d\}$.  After choosing
action $a$, the controller is required to play action $b$ in the next round.
It then reaches a leaf of the tree and continues playing unrestricted.  The
root of the tree is annotated by the initial state $s$ of the abstract game and
the bound $k$ on the number of rounds.  We denote $\textsc{nodes}(T)$ the set
of all nodes of a tree $T$, $\textsc{leaves}(T)$ the subset of leaf nodes.  For
edge $e$, $\textsc{action}(e)$ is the action that labels the edge, and for node
$n$, $\textsc{height}(n)$ is the maximal distance from n to a leaf. $\textsc{height}(T)$ is the height 
of the root node of the tree.  For node
$n$ of the tree, $\textsc{succ}(n)$ is the set of pairs $\langle e, n' \rangle$
where $n'$ is a child node of $n$ and $e$ is the edge connecting $n$ and $n'$.

@@

\begin{algorithm}
    \begin{algorithmic}
        \Function{checkBound}{$k$}
            \State $r \gets $ \Call{solveAbstract}{$\texttt{env}, I, k, \emptyset$}
            \IIf{$r \neq \emptyset$} \Return $r$ \EndIIf
            \State $s' \gets I$
            \For{$x \in X$}
            \State $r \gets$ \Call{solveAbstract}{$\texttt{env}, s' \setminus \{x\}, k, \emptyset$} 
                \IIf{$r = \emptyset$} $s' \gets s' \setminus \{x\}$ \EndIIf \Comment{Removes the assignment to $x$ from $s'$}
            \EndFor
            \State \Return $\emptyset$
        \EndFunction
    \end{algorithmic}

    \caption{Generalise $I$ optimisation}
    \label{alg:opt1}
\end{algorithm}

\subsection{Optimisation: Encouraging loops}

This optimisation is designed to encourage the system player to choose a
candidate strategy that contains loops. If the controller's strategy forces the
game into an error free loop, then the system wins the game. This optimisation
does not attempt to find such a strategy but instead optimistically searches
for candidate strategies that may loop. The modifications add formulas to the
SAT call when the controller is playing. These formulas represent loops of a
particular length in a branch of the tree. The SAT solver is called multiple
times for each branch in the game tree. The search begins by looking for a loop
in the first two game rounds of the branch and successively relaxing this
condition to longer loop lengths. If a loop is found or we reach the end of the
branch then the algorithm moves on to greedily find more loops in other
branches.

\begin{algorithm}
    \begin{algorithmic}
        \Function{loopFmls}{$T$}
            \State $i \gets 0$
            \For{$l \in leaves(gt)$}
                \State $j \gets 0$
                \For{$n \in$ \Call{parents}{$l$}}
                    \For{$p \in$ \Call{parents}{$n$}}
                        \State $loopFmls[i][j] \gets X_n \equiv X_p$
                    \EndFor
                    \State $j \gets j + 1$
                \EndFor
                \State $i \gets i + 1$
            \EndFor
            \State \Return $loopFmls$
        \EndFunction
    \end{algorithmic}

    \begin{algorithmic}
        \Function{findCandidate}{$p, s, k, T$}
        \State $\hat{T} \gets $ \Call{extend}{$T$} \Comment{Extend the tree with unfixed actions}
            \State $f \gets $ \IfElse{$p = \texttt{cont}$}{\Call{treeFormula}{$\hat{T}$}}{\Call{\textoverline{treeFormula}}{$\hat{T}$}} \EndIfElse
            \If{$p = \texttt{cont}$}
                \State $loopFmls \gets $ \Call{loopFmls}{$\hat{T}$}
                \State $currentLoops \gets \top$
                \For{$i \in loopFmls$}
                    \For{$j \in loopFmls[i]$}
                        \State $sol \gets $ \Call{SAT}{$loopFmls[i][j] \land currentLoops \land s(X_{\hat{T}}) \land f$}
                        \If{$sol = \texttt{sat}$}
                            \State $currentLoops \gets currentLoops \land loopFmls[i][j]$
                            \State \textbf{break}
                        \EndIf
                    \EndFor
                \EndFor
            \Else
                \State $sol \gets $ \Call{SAT}{$s(X_{\hat{T}}) \land f$}
            \EndIf
            \If{$sol = \texttt{unsat}$} 
                \If{\texttt{unbounded}} \Comment{Active only in the unbounded solver}
                    %\State $\sigma \gets $ \Call{generalise}{$s$} \Comment{Expand $s$ to a set of states}
                    \State \IfElse{$p = \texttt{cont}$}{\Call{learn}{$s, \hat{T}$}}{\Call{\textoverline{learn}}{$s, \hat{T}$}} \EndIfElse
                \EndIf
                \State \Return $\emptyset$ \Comment{No candidate exists}
            \Else
                \State \Return $\{ \langle n, c \rangle | n \in $ \Call{nodes}{$T$} $, c = \Call{sol}{n} \}$ \Comment{Fix candidate moves in $T$}
            \EndIf
        \EndFunction
    \end{algorithmic}

    \caption{Loop Encouragement}
    \label{alg:opt2}
\end{algorithm}
